Detecting Hate Speech with GPT-3

### 2021-11-04 outputs/documents/paper.Rmd
Update paper will all of Monicas changes and also address that typo and add that paper that we didnt cite

---
output: 
  bookdown::pdf_document2:
    citation_package: natbib
    keep_tex: false
    toc: false
    fig_caption: true
    latex_engine: xelatex
    template: templates/svm-latex-ms.tex
bibliography: "references.bib"
header-includes:
  -  \usepackage{hyperref}
  -  \usepackage{amsmath}
biblio-style: apalike
title: "Detecting Hate Speech with GPT-3"
thanks: "Code and data are available at: https://github.com/kelichiu/GPT3-hate-speech-detection. We gratefully acknowledge the support of Gillian Hadfield and the Schwartz Reisman Institute for Technology and Society. We thank Amy Farrow, Haoluan Chen, Mauricio Vargas Sepúlveda, and Tom Davidson for helpful suggestions. Comments on the `r format(Sys.time(), '%d %B %Y')` version of this paper are welcome at: rohan.alexander@utoronto.ca."
author:
- name: Ke-Li Chiu
  affiliation: University of Toronto
- name: Rohan Alexander
  affiliation: University of Toronto and Schwartz Reisman Institute
abstract: "Sophisticated language models such as OpenAI's GPT-3 can generate hateful text that targets marginalized groups. Given this capacity, we are interested in whether large language models can be used to identify hate speech and classify text as sexist or racist? We use GPT-3 to identify sexist and racist text passages with zero-, one-, and few-shot learning. We find that with zero- and one-shot learning, GPT-3 is able to identify sexist or racist text with an accuracy between 48 per cent and 69 per cent. With few-shot learning and an instruction included in the prompt, the model's accuracy can be as high as 78 per cent. We conclude that large language models have a role to play in hate speech detection, and that with further development language models could be used to counter hate speech and even self-police."
keywords: "GPT-3; natural language processing; quantitative analysis; hate speech."
date: "`r format(Sys.time(), '%B %d, %Y')`"
geometry: margin=1in
fontfamily: mathpazo
fontsize: 12pt
endnote: no
graphics: yes
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(tidy.opts=list(width.cutoff=80), echo = FALSE, warning = FALSE)
# library(dplyr)
library(tidyverse)
# library(tidyr)
# library(readr)
```

@@

Sophisticated natural language processing (NLP) models, such as OpenAI's GPT-3, can produce hateful text. In particular, there have been many examples of text being generated that targets marginalized groups based on their sex, race, sexual orientation, and other characteristics. Large language models are trained on enormous datasets from various sources. This means that untruthful statements, human biases, and abusive language are inevitably included. Further, @bender2021dangers describes how even when the models do not possess intent, they risk producing synthetic texts that are offensive or discriminatory and thus cause unpleasant, or even triggering, interaction experiences.

The sources of the training datasets raise concerns around three issues: exclusion, over-generalization, and exposure [@hovy2016social]. Exclusion happens due to the demographic bias in the dataset. In the case of language models that are trained on US and UK English scraped from the Internet, datasets may be disproportionately white, male, and young. Therefore, it is not surprising to see white supremacist, misogynistic, and ageist content being over-represented in training datasets [@bender2021dangers]. Over-generalization stems from the assumption that what we see in the dataset represents what occurs in the real world. Words such as 'always', 'never', 'everybody', or 'nobody' are frequently used for rhetorical purpose instead of their literal meanings. However, language models do not always recognize this and make inference based on generalized statements using these words. For instance, hate speech commonly uses generalized language for targeting a group such as 'all' and 'every', and a model trained on these statements may generate similarly generalized and harmful statements. Finally, exposure refers to the relative attention, and hence considerations of importance, given to something. In the context of NLP this may be reflected in the emphasis on English-language created under particular circumstances, rather than another languages or circumstances that may be more prevalent.

@@

Even though large NLP models do not have a real understanding of language, the vocabularies and the construction patterns of harmful languages can be thought of as known to them. We show that this knowledge can be used to identify abusive language and even hate speech. In particular we consider 120 different extracts that have been categorized as 'racist', 'sexist', or 'neither' in the single-category settings and 240 different extracts in the mixed-category settings. We ask GPT-3 to classify these based on zero-, one-, and few-shot learning, with and without instruction. We find that the model performs best with few-shot learning when an instruction is included. In that setting the model is able to accurately classify around 78 per cent of the extracts. If language models can be used to identify abusive language, then not only is there potential for them to counter the production of abusive language by humans, but they could also potentially self-police.

@@

In its simplest form, a language model is a probability distribution over sequences of words, which are usually generalized further to tokens. The sequence of tokens constitutes different linguistic units — words, sentences, and even documents [@bengio2003neural]. Language models predict the next token based on inputs. If we consider each token in a vocabulary as a dimension, then the dimensionality of language quickly becomes large [@rosenfeld2000two]. Over time a variety of statistical language models have been created to nonetheless enable prediction. The n-gram is one of the earliest language models. It works by considering the co-occurrence of tokens in a sequence. In the early 2000s, prominent neural network language models were developed, for instance @bengio2003neural. These were then built on by word embeddings language models in the 2010s in which the distance between tokens represents how related those tokens are, for instance @turian2010word. In 2017, @vaswani2017attention introduced the Transformer, which marked a new era for language models. The Transformer is a network architecture for neural networks that can be trained more quickly than many other approaches [@vaswani2017attention]. Now most of the representative pre-trained language models, such as BERT [@devlin2018bert], GPT-2 [@radford2019language], and GPT-3 [@brown2020language], are built on this architecture.

GPT-3 is the third generation of the Generative Pre-trained Transformer models created by OpenAI. Until January 11, when @fedus2021switch announced a Transformer model with a trillion parameters, GPT-3 was the largest, publicly-known, Transformer language model. GPT-3 is distinctive from its predecessors because of few-shot learning. This means that GPT-3 can 'learn' to perform a new task based on only a few examples, expressed in natural language, instead of a fine-tuning process that can require a large amount of data. GPT-3 has led to unexpected NLP applications, such as computational code generation given natural language prompts. However, like other language models, GPT-3 has also generated inappropriate or even hateful content. For instance, @mcguffie2020radicalization demonstrated the use of GPT-3 in mass-producing radicalized text targeting the Islamic population.

@@

Hate speech detection is of interest to researchers in a variety of domains including computer science and sociology, as well as industry and the judiciary. Detecting hate speech is difficult because the definition of hate speech varies depending on the complex intersection of the topic of the assertion, the context, the timing of the post, synchronized world events, and the identity of speaker and recipient [@schmidt2017survey]. Moreover, it is difficult to discern hate speech from merely offensive language [@davidson2017automated]. Since hate speech is prohibited in several countries, misclassification of hate speech can become a legal problem. For instance, in Canada, speech that contains 'public incitement of hatred' or 'wilful promotion of hatred' is specified by the Criminal Code [@act2021justice]. Policies toward hate speech are more detailed in some social media platforms. For instance, the Twitter Hateful Conduct Policy states:

@@

There has been a large amount of research focused on detecting hate speech. And as part of this various hate speech datasets have been created and examined. For instance, @waseem2016hateful create a dataset that captures hate speech in the form of racist and sexist language that includes domain expert annotation, and @davidson2017automated trains a classifier to distinguish between hate speech and offensive language. And it is important to note that even these datasets have bias. For instance, @davidson2019racial found racial bias in five different sets of Twitter data annotated for hate speech and abusive language. They found that tweets written in African American English are more likely to be labeled as abusive.

@@

We examine the ability of GPT-3 to identity hate speech in zero-shot, one-shot, and few-shot settings. There are a variety of parameters, such as temperature, that control the degree of text variation. To enhance consistency, the temperature is set to zero in our experiments. There are two categories of hate speech that are of interest in this paper. The first targets the race of the recipient, and the other targets the gender of the recipient. With zero- and one-shot learning, the model identifies hate speech one category at a time. With few-shot learning, the categories are mixed, and the model is asked to classify an input as sexist, racist, or neither.

@@

We use the ETHOS dataset created by @mollas2020ethos. ETHOS is based on comments found in YouTube and Reddit. The ETHOS YouTube data is collected through Hatebusters [@anagnostou2018hatebusters]. Hatebusters is a platform that collects comments from YouTube and assigns a 'hate' score to them using a support vector machine. The Reddit data is collected from the Public Reddit Data Repository [@baumgartner2020pushshift]. The dataset has two variants: binary and multi-label. In the binary dataset comments are classified as hate or non-hate. In the multi-label variant, the comments are evaluated on measures that include violence, gender, race, ability, religion, and sexual orientation. The examples provided in this paper are from the ETHOS dataset and hence contain typos, misspelling, and offensive content.

There are 998 statements in the ETHOS dataset that have a binary classification of hate speech or not hate speech. Of these, the 433 statements that contain hate speech additionally have multiple labels that classify the content, for instance, does the comment have to do with violence, gender, race, nationality, disability, etc. We considered the 136 statements that contain race-based hate speech, and we focus on the 76 whose race-based score is at least 0.5. Similarly, we considered the 174 statements that contain gender-based hate speech, and again focused on the 84 whose gender-based score is at least 0.5. We randomly draw 120 of these statements, using stratified sampling, to create a balanced dataset of 60 sexist statements and 60 racist statements. For comparison purposes, we randomly draw 120 statements that are classified as not hate speech.

@@

@@

> 'Too bad women don’t know how to kill themselves': sexist.
>
> 'You should use your time to arrest murderers not little kids': not-sexist.
> 
> 'Now they know better than this shit lol they dudes. The stronger sex. The man supremacy': sexist.
> 
> 'The thing is women are not equal to us men and their place is the home and kitchen:'

@@

@@

> 'That girl still need a good ass whooping, stupid ass bitch!!': sexist. 
>
> 'U SUCK SHIT YOU BLACK FUCK': racist. 
>
> 'You deserve way more subscribers. The shit you say is fucking gold.': neither. 
>
> 'Every girl on this planet curses her mother for their birth. They know that they are the powerless sex and they pray that they had been born as men! With dicks and stuff.':

@@

> Classify the following texts into 'racist', 'sexist' or 'neither'. 
>
> 'That girl still need a good ass whooping, stupid ass bitch!!': sexist. 
>
> 'U SUCK SHIT YOU BLACK FUCK': racist. 
>
> 'You deserve way more subscribers. The shit you say is fucking gold.': neither. 
>
> 'Every girl on this planet curses her mother for their birth. They know that they are the powerless sex and they pray that they had been born as men! With dicks and stuff.':

@@

The results of the zero-shot experiments are presented in Table \@ref(tab:zeroshot). The model has 35 matches and 25 mismatches in the sexist category, and 23 matches and 37 mismatches in the racist category. The model performs better when identifying sexist comments compared with identifying racist comments. However, the overall ratio of matches and mismatches is 58:62. In other words, the accuracy in identifying hate speech in the zero-shot setting is 48.3 per cent.

```{r zeroshot}
zero_shot_result <- read.csv(here::here("outputs/data/zero_shot_results.csv"))

zero_shot_result <- 
  zero_shot_result %>%
  filter(temperature == 0) %>%
  select(category, label, answer, temperature) %>%
  mutate(label_yn = ifelse(grepl("not", label), "N", "Y")) %>%
  mutate(answer_yn = ifelse(stringr::str_detect(answer, "No", negate = FALSE), "N", "Y")) %>%
  mutate(result = case_when(answer_yn==label_yn ~ "Match",
                            answer_yn!=label_yn ~ "Mismatch"
                            )) %>%
  group_by(result, category) %>%
  summarise(result_cnt = n(), .groups="drop") %>% 
  mutate(category = case_when(category == "racist" ~ "Racist",
                              category == "sexist" ~ "Sexist"
                              )) %>%
  rename(Result = result,
         Category = category,
         Count = result_cnt)

@@

one_shot_result <- 
  one_shot_result %>%
  filter(temperature == 0) %>%
  select(category, label, answer) %>%
  mutate(label_yn = ifelse(grepl("not", label), "N", "Y")) %>%
  mutate(answer_yn = ifelse(stringr::str_detect(answer, "No", negate = FALSE), "N", "Y")) %>%
  mutate(result = case_when(answer_yn==label_yn ~ "Match",
                            answer_yn!=label_yn ~ "Mismatch"
                            )) %>%
  group_by(result, category) %>%
  summarise(result_cnt = n(), .groups = "drop") %>% 
  mutate(category = case_when(category == "racist" ~ "Racist",
                              category == "sexist" ~ "Sexist"
                              )) %>%
  rename(Result = result,
         Category = category,
         Count = result_cnt)

@@

few_shot_single_result <- 
  few_shot_single_result %>%
  filter(temperature == 0) %>%
  select(category, label, answer) %>%
  mutate(label_yn = ifelse(grepl("not", label), "No", "Yes")) %>%
  mutate(answer_yn = ifelse(grepl("not", answer), "No", "Yes")) %>%
  mutate(result = case_when(answer_yn==label_yn ~ "Match",
                            answer_yn!=label_yn ~ "Mismatch"
                            )) %>%
  mutate(result_with_cat = case_when(answer_yn==label_yn & answer_yn == "Yes" ~ "True Positive",
                                     answer_yn==label_yn & answer_yn == "No"  ~ "True Negative",
                                     answer_yn!=label_yn & answer_yn == "Yes" ~ "False Positive",
                                     answer_yn!=label_yn & answer_yn == "No"  ~ "False Negative"
                                     )) %>%
  group_by(result, category) %>%
  summarise(result_cnt = n(), .groups = "drop") %>% 
  mutate(category = case_when(category == "racist" ~ "Racist",
                              category == "sexist" ~ "Sexist"
                              )) %>%
  rename(Result = result,
         Category = category,
         Count = result_cnt)

@@

few_shot_mixed_result <- 
  few_shot_mixed_result %>%
  filter(temperature == 0) %>%
  select(category, label, answer, example_set) %>%
  mutate(example_set = example_set+1) %>%
  mutate(result = ifelse(trimws(label)==trimws(answer), "Match", "Mismatch")) %>%
  mutate(label = case_when(label == "racist" ~ "Racist",
                           label == "sexist" ~ "Sexist",
                           label == "neither" ~ "Neither"
                           )) %>%
  group_by(example_set, result) %>%#, label) %>%
  summarise(result_cnt = n(), .groups = "drop") %>% 
  rename(Result = result,
         "Example Set" = example_set,
         Count = result_cnt)

knitr::kable(few_shot_mixed_result, 
             caption = "Classification of statements with mixed-category few-shot learning", 
             booktabs = TRUE,
             linesep = "",
             align = 'llr')
```

The unique generated answers are listed in Table \@ref(tab:fewshotmixedanswersnoinstruct). Under the mixed-category setting, the model is observed to generate answers that are out of scope. For instance, other than 'sexist', 'racist', and 'neither', we also see answers such as 'transphobic', 'hypocritical', 'Islamophobic', and 'ableist'. In some cases, the model even classifies a text passage into more than one category, such as 'sexist, racist' and 'sexist, misogynistic'. The full list contains eighty different answers instead of three.

@@

few_shot_mixed_result <- few_shot_mixed_result %>% 
  select(answer) %>% 
  distinct() %>% 
  rename(Answer = answer)# %>% 
few_shot_mixed_result <-list(few_shot_mixed_result$Answer)

df <- as.data.frame("Answer")
df <- rbind(df$Answer,paste(unlist(few_shot_mixed_result[[1]]), collapse=' | '))
df %>% 
  knitr::kable(
  caption = "Classifications generated by GPT-3 under mixed-category few-shot learning without instructions",
  booktabs = TRUE,
  linesep = "") %>% 
  kableExtra::column_spec(column = 1, width = "40em")
```

@@

To reduce the chance of the model generating answers that are out of scope, a brief instruction is added to the prompt, specifying that the answers be: 'sexist', 'racist', or 'neither'. The addition of an instruction successfully restricts the generated answers within the specified terms. The unique generated answers are listed in Table \@ref(tab:fewshotmixedanswerswithinstruct).

```{r fewshotmixedanswerswithinstruct, message=FALSE, warning=FALSE}
few_shot_instruction_result <-
  readr::read_csv(here::here("outputs/data/few_shot_fixed_examples_instruction_results.csv"))

few_shot_instruction_result %>% 
  select(answer) %>% 
  distinct() %>% 
  rename(Answer = answer) %>% 
  knitr::kable(caption = "Classifications generated by GPT-3 under mixed-category few-shot learning with instructions", 
             booktabs = TRUE,
             linesep = "",
             align = 'llr')
```

>>>>

---
output: 
  bookdown::pdf_document2:
    citation_package: natbib
    keep_tex: false
    toc: false
    fig_caption: true
    latex_engine: xelatex
    template: templates/svm-latex-ms.tex
bibliography: "references.bib"
header-includes:
  -  \usepackage{hyperref}
  -  \usepackage{amsmath}
biblio-style: apalike
title: "Detecting Hate Speech with GPT-3"
thanks: "Code and data are available at: https://github.com/kelichiu/GPT3-hate-speech-detection. We gratefully acknowledge the support of Gillian Hadfield, the Schwartz Reisman Institute for Technology and Society, and OpenAI for providing access to GPT-3 under the academic access program. We thank Amy Farrow, Haoluan Chen, John Giorgi, Mauricio Vargas Sepúlveda, Monica Alexander, Noam Kolt, and Tom Davidson for helpful discussions and suggestions. Please note that we have added asterisks to racial slurs and other offensive content in this paper, however the inputs and outputs did not have these.  Comments on the `r format(Sys.time(), '%d %B %Y')` version of this paper are welcome at: rohan.alexander@utoronto.ca."
author:
- name: Ke-Li Chiu
  affiliation: University of Toronto
- name: Rohan Alexander
  affiliation: University of Toronto and Schwartz Reisman Institute
abstract: "Sophisticated language models such as OpenAI's GPT-3 can generate hateful text that targets marginalized groups. Given this capacity, we are interested in whether large language models can be used to identify hate speech and classify text as sexist or racist? We use GPT-3 to identify sexist and racist text passages with zero-, one-, and few-shot learning. We find that with zero- and one-shot learning, GPT-3 can identify sexist or racist text with an accuracy between 48 per cent and 69 per cent. With few-shot learning and an instruction included in the prompt, the model's accuracy can be as high as 78 per cent. We conclude that large language models have a role to play in hate speech detection, and that with further development language models could be used to counter hate speech and even self-police."
keywords: "GPT-3; natural language processing; quantitative analysis; hate speech."
date: "`r format(Sys.time(), '%B %d, %Y')`"
geometry: margin=1in
fontfamily: mathpazo
fontsize: 12pt
endnote: no
graphics: yes
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(tidy.opts=list(width.cutoff=80), echo = FALSE, warning = FALSE)
# library(dplyr)
library(tidyverse)
# library(tidyr)
# library(readr)
```

@@

**This paper contains language and themes that are offensive.**

Natural language processing (NLP) models focus on using normal words, often written text, as their data. For instance, one might have content from a large number of books and want to group them into themes. Sophisticated NLP models are being increasingly embedded in society. For instance, Google Search uses an NLP model called BERT to better understand what is meant by a word given its context. Some sophisticated natural language processing (NLP) models, such as OpenAI's GPT-3, can additionally produce text as an output.

The text produced by sophisticated NLP models can be hateful. In particular, there have been many examples of text being generated that targets marginalized groups based on their sex, race, sexual orientation, and other characteristics. For instance, 'Tay' was a chatbot that was released by Microsoft in 2016 on Twitter. Within hours of being released, some of its tweets were sexist. Large language models are trained on enormous datasets from various sources. This means that untruthful statements, human biases, and abusive language are inevitably included. Further, even when the models do not possess intent, they risk producing synthetic texts that are offensive or discriminatory and thus cause unpleasant, or even triggering, interaction experiences [@bender2021dangers].

Often the datasets that underpin these models consist of, essentially, the whole public internet. This source raises concerns around three issues: exclusion, over-generalization, and exposure [@hovy2016social]. Exclusion happens due to the demographic bias in the dataset. In the case of language models that are trained on US and UK English scraped from the Internet, datasets may be disproportionately white, male, and young. Therefore, it is not surprising to see white supremacist, misogynistic, and ageist content being over-represented in training datasets [@bender2021dangers]. Over-generalization stems from the assumption that what we see in the dataset represents what occurs in the real world. Words such as 'always', 'never', 'everybody', or 'nobody' are frequently used for rhetorical purpose instead of their literal meanings. However, language models do not always recognize this and make inference based on generalized statements using these words. For instance, hate speech commonly uses generalized language for targeting a group such as 'all' and 'every', and a model trained on these statements may generate similarly generalized and harmful statements. Finally, exposure refers to the relative attention, and hence considerations of importance, given to something. In the context of NLP this may be reflected in the emphasis on English-language created under particular circumstances, rather than another languages or circumstances that may be more prevalent.

@@

Even though large NLP models do not have a real understanding of language, the vocabularies and the construction patterns of harmful languages can be thought of as known to them. We show that this knowledge can be used to identify abusive language and even hate speech. In particular we consider 120 different extracts that have been categorized as 'racist', 'sexist', or 'neither' in the single-category settings and 240 different extracts in the mixed-category settings. We ask GPT-3 to classify these based on zero-, one-, and few-shot learning, with and without instruction. We find that the model performs best with few-shot learning when an instruction is included. In that setting the model can accurately classify around 78 per cent of the extracts. If language models can be used to identify abusive language, then not only is there potential for them to counter the production of abusive language by humans, but they could also potentially self-police.

@@

In its simplest form, a language model involves assigning a probability to a certain sequence of words. For instance, the sequence 'the cat in the hat' is probably more likely than 'the cat in the computer'. We typically talk of tokens, or collections of characters, rather than words, and a sequence of tokens constitutes different linguistic units — words, sentences, and even documents [@bengio2003neural]. Language models predict the next token based on inputs. If we consider each token in a vocabulary as a dimension, then the dimensionality of language quickly becomes large [@rosenfeld2000two]. Over time a variety of statistical language models have been created to nonetheless enable prediction. The n-gram is one of the earliest language models. It works by considering the co-occurrence of tokens in a sequence. For instance, given the four-word sequence, 'the cat in the', it is more likely that the fifth word is 'hat' rather than 'computer'. In the early 2000s, language models based on neural networks were developed, for instance @bengio2003neural. These were then built on by word embeddings language models in the 2010s in which the distance between tokens represents how related those tokens are, for instance @turian2010word. In 2017, @vaswani2017attention introduced the Transformer, which marked a new era for language models. The Transformer is a network architecture for neural networks that can be trained more quickly than many other approaches [@vaswani2017attention]. Now most representative pre-trained language models, such as Google's model: Bidirectional Encoder Representations from Transformers (BERT) [@devlin2018bert], as well as OpenAI's: Generative Pre-trained Transformer (GPT)-2 [@radford2019language], and GPT-3 [@brown2020language], are built on this architecture. These models are widely used, for instance BERT is used by Google search.

GPT-3 is the third generation of the Generative Pre-trained Transformer models created by OpenAI. OpenAI is a private company in California that develops artificial intelligence models. Until January 11, when @fedus2021switch announced a Transformer model with a trillion parameters, GPT-3 was the largest, publicly-known, Transformer language model. GPT-3 is distinctive from its predecessors because of few-shot learning. This means that GPT-3 can 'learn' to perform a new task based on only a few examples, expressed in natural language, instead of a fine-tuning process that can require a large amount of data. GPT-3 has led to unexpected NLP applications, such as computational code generation given natural language prompts.

Like other language models, GPT-3 has also generated inappropriate or even hateful content. For instance, @mcguffie2020radicalization demonstrated the use of GPT-3 in mass-producing radicalized text targeting the Islamic population. And @lin2021truthfulqa show that GPT-3 and similar language models can propagate misconceptions that could deceive human readers. For instance, when asked 'Who really caused 9/11?', they found that GPT-3 provided the false statement 'The US government caused 9/11.'

@@

There is no commonly held definition of hate speech and different legal jurisdictions have different definitions, as do different companies and other groups. One definition is 'the intentional verbalization of prejudice against a social group' [@kennedy2018gab]. Detecting hate speech is difficult because the definition of hate speech varies depending on the complex intersection of the topic of the assertion, the context, the timing of the post, synchronized world events, and the identity of speaker and recipient [@schmidt2017survey]. Moreover, it is difficult to discern hate speech from merely offensive language [@davidson2017automated]. Hate speech detection is of interest to academic researchers in a variety of domains including computer science [@addressinghatespeech] and sociology [@davidson2017automated]. It is also of interest to industry, for instance to maintain standards on social networks, and in the judiciary to help identify and prosecute crimes.  Since hate speech is prohibited in several countries, misclassification of hate speech can become a legal problem. For instance, in Canada, speech that contains 'public incitement of hatred' or 'wilful promotion of hatred' is specified by the Criminal Code [@act2021justice]. Policies toward hate speech are more detailed in some social media platforms. For instance, the Twitter Hateful Conduct Policy states:

@@

There has been a large amount of research focused on detecting hate speech. And as part of this various hate speech datasets have been created and examined. For instance, @waseem2016hateful create a dataset that captures hate speech in the form of racist and sexist language that includes domain expert annotation. The use Twitter data, and annotate 16,914 tweets: 3,383 as sexist, 1,972 as racist, and 11,559 as neither. They had a high degree of annotator agreement, and most of the disagreements were to do with sexism, and often explained by an annotator lacking apparent context. And, @davidson2017automated trains a classifier to distinguish between hate speech and offensive language. To define hate speech, they use online 'hate speech lexicon containing words and phrases identified by internet users as hate speech'. It is important to note that even these datasets have bias. For instance, @davidson2019racial found racial bias in five different sets of Twitter data annotated for hate speech and abusive language. They found that tweets written in African American English are more likely to be labeled as abusive.

@@

We examine the ability of GPT-3 to identity hate speech in zero-shot, one-shot, and few-shot settings. There are a variety of parameters, such as temperature, that control the degree of text variation. Temperature is a hyper-parameter between zero and one. Lower temperatures mean that the model places more weight on higher-probability tokens. To enhance consistency, the temperature is set to zero in our experiments and not varied. There are two categories of hate speech that are of interest in this paper. The first targets the race of the recipient, and the other targets the gender of the recipient. With zero- and one-shot learning, the model identifies hate speech one category at a time. With few-shot learning, the categories are mixed, and the model is asked to classify an input as sexist, racist, or neither. Zero-shot learning means an example is not provided in the prompt. One-shot learning means that one example is provided, and few-shot means that two or more examples are provided.

@@

We use the ETHOS dataset created by @mollas2020ethos. ETHOS is based on comments found in YouTube and Reddit. The ETHOS YouTube data is collected through Hatebusters [@anagnostou2018hatebusters]. Hatebusters is a platform that collects comments from YouTube and assigns a 'hate' score to them using a support vector machine. That hate-score is only used to decide whether to consider the comment further or not. The Reddit data is collected from the Public Reddit Data Repository [@baumgartner2020pushshift]. The classification is done by contributors to a crowd-sourcing platform. They are first asked with an example contains hate speech, and then, if it does, whether it incites violence and other additional details. The dataset has two variants: binary and multi-label. In the binary dataset comments are classified as hate or non-hate based. In the multi-label variant, the comments are evaluated on measures that include violence, gender, race, ability, religion, and sexual orientation. The examples provided in this paper are from the ETHOS dataset and hence contain typos, misspelling, and offensive content.

We use all of the 998 statements in the ETHOS dataset that have a binary classification of hate speech or not hate speech. Of these, the 433 statements that contain hate speech additionally have multiple labels that classify the content, for instance, does the comment have to do with violence, gender, race, nationality, disability, etc. We considered all of the 136 statements that contain race-based hate speech, and we focus on the 76 whose race-based score is at least 0.5. Similarly, we considered the all of 174 statements that contain gender-based hate speech, and again focused on the 84 whose gender-based score is at least 0.5. To create a balanced dataset we randomly draw 120 of these statements, using stratified sampling, to result in 60 sexist statements and 60 racist statements. For comparison purposes, we randomly draw 120 statements that are classified as not hate speech. The balanced sample enables us to more easily understand our results.

@@

The model provides this answer based on the dataset that it was trained on.

@@

> 'Too bad women don’t know how to kill themselves': sexist.
>
> 'You should use your time to arrest murderers not little kids': not-sexist.
> 
> 'Now they know better than this s*** lol they dudes. The stronger sex. The man supremacy': sexist.
> 
> 'The thing is women are not equal to us men and their place is the home and kitchen:'

@@

The assignment of sexist or not in the prompt is based on the ETHOS dataset classification.

@@

> 'That girl still need a good ass whooping, stupid ass b****!!': sexist. 
>
> 'U SUCK S*** YOU BLACK F***': racist. 
>
> 'You deserve way more subscribers. The s*** you say is f***ing gold.': neither. 
>
> 'Every girl on this planet curses her mother for their birth. They know that they are the powerless sex and they pray that they had been born as men! With d***s and stuff.':

@@

> Classify the following texts into 'racist', 'sexist' or 'neither'. 
>
> 'That girl still need a good ass whooping, stupid ass b****!!': sexist. 
>
> 'U SUCK S*** YOU BLACK F***': racist. 
>
> 'You deserve way more subscribers. The s*** you say is f***ing gold.': neither. 
>
> 'Every girl on this planet curses her mother for their birth. They know that they are the powerless sex and they pray that they had been born as men! With d***s and stuff.':

@@

The results of the zero-shot experiments are presented in Table \@ref(tab:zeroshot). The model has 35 matches and 25 mismatches in the sexist category, and 23 matches and 37 mismatches in the racist category. The model performs better when identifying sexist comments compared with identifying racist comments. However, the overall ratio of matches and mismatches is 58:62. In other words, the accuracy in identifying hate speech in the zero-shot setting is 48 per cent.

```{r zeroshot}
zero_shot_result <- read_csv(here::here("outputs/data/zero_shot_results.csv"), show_col_types = FALSE)

zero_shot_result <- 
  zero_shot_result %>%
  filter(temperature == 0) %>%
  select(category, label, answer, temperature) %>%
  mutate(label_yn = ifelse(grepl("not", label), "N", "Y")) %>%
  mutate(answer_yn = ifelse(stringr::str_detect(answer, "No", negate = FALSE), "N", "Y")) %>%
  mutate(result = case_when(answer_yn==label_yn ~ "Match",
                            answer_yn!=label_yn ~ "Mismatch"
                            )) %>%
  group_by(result, category) %>%
  summarise(result_cnt = n(), .groups="drop") %>% 
  mutate(category = case_when(category == "racist" ~ "Racist",
                              category == "sexist" ~ "Sexist"
                              )) %>%
  mutate(Percent = result_cnt / sum(result_cnt) * 100,
         Percent = round(Percent)) %>%
  rename(Result = result,
         Category = category,
         Count = result_cnt)

@@

one_shot_result <- 
  one_shot_result %>%
  filter(temperature == 0) %>%
  select(category, label, answer) %>%
  mutate(label_yn = ifelse(grepl("not", label), "N", "Y")) %>%
  mutate(answer_yn = ifelse(stringr::str_detect(answer, "No", negate = FALSE), "N", "Y")) %>%
  mutate(result = case_when(answer_yn==label_yn ~ "Match",
                            answer_yn!=label_yn ~ "Mismatch"
                            )) %>%
  group_by(result, category) %>%
  summarise(result_cnt = n(), .groups = "drop") %>% 
  mutate(category = case_when(category == "racist" ~ "Racist",
                              category == "sexist" ~ "Sexist"
                              )) %>%
  mutate(Percent = result_cnt / sum(result_cnt) * 100,
         Percent = round(Percent)) %>%
  rename(Result = result,
         Category = category,
         Count = result_cnt)

@@

few_shot_single_result <- 
  few_shot_single_result %>%
  filter(temperature == 0) %>%
  select(category, label, answer) %>%
  mutate(label_yn = ifelse(grepl("not", label), "No", "Yes")) %>%
  mutate(answer_yn = ifelse(grepl("not", answer), "No", "Yes")) %>%
  mutate(result = case_when(answer_yn==label_yn ~ "Match",
                            answer_yn!=label_yn ~ "Mismatch"
                            )) %>%
  mutate(result_with_cat = case_when(answer_yn==label_yn & answer_yn == "Yes" ~ "True Positive",
                                     answer_yn==label_yn & answer_yn == "No"  ~ "True Negative",
                                     answer_yn!=label_yn & answer_yn == "Yes" ~ "False Positive",
                                     answer_yn!=label_yn & answer_yn == "No"  ~ "False Negative"
                                     )) %>%
  group_by(result, category) %>%
  summarise(result_cnt = n(), .groups = "drop") %>% 
  mutate(category = case_when(category == "racist" ~ "Racist",
                              category == "sexist" ~ "Sexist"
                              )) %>%
  mutate(Percent = result_cnt / sum(result_cnt) * 100,
         Percent = round(Percent)) %>%
  rename(Result = result,
         Category = category,
         Count = result_cnt)

@@

few_shot_mixed_result <-
  few_shot_mixed_result %>%
  filter(temperature == 0) %>%
  select(category, label, answer, example_set) %>%
  mutate(example_set = example_set + 1) %>%
  mutate(result = ifelse(trimws(label) == trimws(answer), "Match", "Mismatch")) %>%
  mutate(
    label = case_when(
      label == "racist" ~ "Racist",
      label == "sexist" ~ "Sexist",
      label == "neither" ~ "Neither"
    )
  ) %>%
  group_by(example_set, result) %>% #, label) %>%
  summarise(result_cnt = n(), .groups = "drop") %>%
  ungroup() %>%
  group_by(example_set) %>%
  mutate(Percent = result_cnt / sum(result_cnt) * 100,
         Percent = round(Percent)) %>%
  rename(Result = result,
         "Example Set" = example_set,
         Count = result_cnt)

knitr::kable(
  few_shot_mixed_result,
  caption = "Classification of statements with mixed-category few-shot learning",
  booktabs = TRUE,
  linesep = "",
  align = 'lllr'
)
```

The unique generated answers are listed in Table \@ref(tab:fewshotmixedanswersnoinstruct). These are the response of GPT-3 that we obtain when we ask the model to classify statements, but do not provide examples that would serve to limit the responses. Under the mixed-category setting, the model is observed to generate answers that are out of scope. For instance, other than 'sexist', 'racist', and 'neither', we also see answers such as 'transphobic', 'hypocritical', 'Islamophobic', and 'ableist'. In some cases, the model even classifies a text passage into more than one category, such as 'sexist, racist' and 'sexist, misogynistic'. The full list contains eighty different answers instead of three.

@@

few_shot_mixed_result <- few_shot_mixed_result %>%
  select(answer) %>%
  distinct() %>%
  rename(Answer = answer)# %>%
few_shot_mixed_result <- list(few_shot_mixed_result$Answer)

df <- as.data.frame("Answer")
df <-
  rbind(df$Answer, paste(unlist(few_shot_mixed_result[[1]]), collapse = ' | '))
df %>%
  knitr::kable(caption = "Classifications generated by GPT-3 under mixed-category few-shot learning without instructions",
               booktabs = TRUE,
               linesep = "") %>%
  kableExtra::column_spec(column = 1, width = "40em")
```

@@

To reduce the chance of the model generating answers that are out of scope, a brief instruction is added to the prompt, specifying that the answers be: 'sexist', 'racist', or 'neither'. The addition of an instruction successfully restricts the generated answers within the specified terms. The unique generated answers are: 'racist', 'neither', and 'sexist'.

```{r fewshotmixedanswerswithinstruct, message=FALSE, warning=FALSE}
# few_shot_instruction_result <-
#   readr::read_csv(here::here("outputs/data/few_shot_fixed_examples_instruction_results.csv"))
# 
# few_shot_instruction_result %>% 
#   select(answer) %>% 
#   distinct()
# %>% 
#   rename(Answer = answer) %>% 
#   knitr::kable(caption = "Classifications generated by GPT-3 under mixed-category few-shot learning with instructions", 
#              booktabs = TRUE,
#              linesep = "",
#              align = 'llr')
```

The results of the mixed-category, few-shot learning, with instruction, experiments are presented in Table \@ref(tab:fewshotmixedinstruct). With the addition of an instruction in the prompt, the example set that yields the best result is Example Set 1 instead of Example Set 2. The addition of the instruction in the prompt increases the highest number of matches from 180 to 187; the highest accuracy rate is increased from 75 per cent to 78 per cent. In almost all cases the accuracy of the model increases (Figure \ref{fig:comparison}). Across all the example sets, there are 1,785, or 74 per cent, matches and 615, or 26 per cent, mismatches.
